R Notebook

### 2017-04-01 doc/RMD/MLG-distribution.Rmd
Rework MLG-graph

```{r, fig.width = 10, fig.height = 10}
make_from_to <- function(poplist){
  x <- combn(poplist, 2)
  tibble::data_frame(from = x[1, ], to = x[2, ])
}
crosses <- mlg.crosspop(dat, df = TRUE, quiet = TRUE)

pg <- 
  crosses %>%
  group_by(MLG) %>% 
  summarize(Population = list(make_from_to(Population)), Size = length(Count))

pop_graph <- pg %>%  unnest() %>%
  group_by(from, to) %>%
  mutate(weight = length(Size)) %>%
  ungroup() %>%
  select(from, to, weight, everything())

unchoose <- function(x) ceiling(sqrt(x * 2))

gcross           <- graph_from_data_frame(pop_graph, directed = FALSE)
E(gcross)$name   <- as.character(pop_graph$MLG)
E(gcross)$weight <- unchoose(table(E(gcross)$name)[E(gcross)$name])
colors <- viridis::magma(max(E(gcross)$weight) - 1, end = 0.8)
set.seed(500)
gcross %>%
  plot(., vertex.size = table(pop(clonecorrect(dat, ~Region)))[names(V(gcross))],
       edge.width = E(.)$weight - 1,
       layout = layout_as_star(., center = "MI"),
       # layout = layout_in_circle(.),
       edge.color = colors[E(.)$weight - 1])

```

```{r, fig.width = 10, fig.height = 10}
psize <- table(pop(clonecorrect(dat, ~Region)))
pops <- crosses %>%
  mutate_if(is.factor, as.character) %>%
  group_by(Population) %>%
  summarize(MLG = list(as.character(MLG))) %>%
  group_by(Population) %>%
  mutate(size = psize[unlist(Population)]) %>%
  mutate(n_private = size - lengths(MLG)) %>%
  mutate(interactions = list(pg$Population[pg$MLG %in% unlist(MLG)] %>% 
                               setNames(unlist(MLG)) %>% 
                               bind_rows(.id = "MLG") %>% 
                               distinct() %>%
                               filter(Population == from | Population == to))) %>%
  mutate(graph = map(interactions, ~select(.x, from, to, MLG) %>% 
                       rename(name = MLG) %>% 
                       graph_from_data_frame(directed = FALSE)))
pops
par(mfrow = c(3, 4))

apply(pops, 1, function(i){
    plot(i$graph, 
         layout = layout_as_star(i$graph, center = i$Population), 
         main = i$Population, 
         vertex.size = psize[names(V(i$graph))])
  })
par(mfrow = c(1, 1))

gnew <- igraph::union(pops$graph[[1]], pops$graph[-1])

glayout <- layout_as_star(gnew, center = "MI")

glayout_df <- glayout %>% 
  data.frame %>% 
  setNames(c("x", "y")) %>% 
  as.list %>% 
  c(Population = list(names(V(gnew)))) %>% 
  dplyr::as_data_frame() %>%
  inner_join(pops)
V(gnew)$size <- glayout_df$size
V(gnew)$color <- viridis::viridis(100)[round(100 * (1 - glayout_df$n_private/glayout_df$size))]
gnew <- add_vertices(gnew, length(V(gnew)), size = glayout_df$n_private, color = "grey90")
glayout2 <- rbind(glayout, glayout)
plot(gnew, layout = glayout2)

```

>>>>

I realized that it's possible to use an MLG table with matrix multiplication to
get an adjency matrix.

```{r}
datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
adjmat  <- datmlg %*% t(datmlg) 
cols    <- sort(colnames(adjmat))
adjmat  <- adjmat[cols, cols]
adjmat
crosses
```

Now that we have the adjacency matrix, we can use it to construct our graph:

```{r, fig.width = 10, fig.height = 10}
g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
V(g)$size   <- diag(adjmat)
g           <- delete_vertices(g, degree(g) == 0)
shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
V(g)$weight <- 1 - shared_mlg/V(g)$size
el          <- as_adj_edge_list(g)
el          <- el[lengths(el) > 0]
popgraphs <- setNames(vector(mode = "list", length = length(el)), names(el))
for (v in names(el)){
  idx  <- el[[v]]
  mlgs <- crosses %>%           # How to get all MLGs from a single population:
    filter(Population == v) %>%         # Grab only the population e and then
    select(MLG) %>%                     # remove everything but the MLGs to do an
    inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
    filter(Population != v) %>%         # remove the query population to give
    arrange(Population)                 # the neigboring populations in order.
  MLGS <- as.character(mlgs$MLG)
  E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
  E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
  popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
}
par(mfrow = c(3, 4))
for (i in names(popgraphs)){
  pg <- popgraphs[[i]]
  labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
  labs <- ifelse(duplicated(labs), NA, labs)
  plot(pg, 
       main = i, 
       layout = layout_as_star(pg, center = i), 
       edge.width = E(pg)$weight,
       edge.label = labs)
}
par(mfrow = c(1, 1))
center_node <- degree(g) %>% which.max() %>% names()

new_layout <- structure(c(1, 0.809016994374947, 0.309016994374947, -0.309016994374947, 
-0.809016994374947, -1, -0.809016994374947, -0.309016994374948, 
0, 0.309016994374947, 0.809016994374947, 0, 0.587785252292473, 
0.951056516295154, 0.951056516295154, 0.587785252292473, 1.22464679914735e-16, 
-0.587785252292473, -0.951056516295154, 0, -0.951056516295154, 
-0.587785252292473), .Dim = c(11L, 2L), .Dimnames = list(c("MN", 
"AU", "OR", "FR", "WA", "CA", "NE", "CO", "MI", "ND", "NY"), 
    c("x", "y")))
new_layout <- new_layout[V(g)$name, ]
g <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
plot(g, 
     layout = rbind(new_layout, new_layout),
     edge.width = E(g)$weight,
     edge.label = NA)

glay <- create_layout(g, "manual", node.positions = as.data.frame(rbind(new_layout, new_layout)))
ggraph(glay) +
  geom_edge_fan(aes(alpha = weight + 1)) +
  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = weight)) +
  geom_node_label(aes(label = name), repel = TRUE) +
  viridis::scale_fill_viridis(option = "C") +
  coord_fixed() +
  theme_void() +
  labs(list(
    title = "Shared haplotypes across regions",
    fill = "Number of\nGenotypes",
    alpha = "Fraction of\nprivate genotpes",
    edge_alpha = "Populations\nper haploytpe",
    caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
  ))
```
